---
title: "Assignment2"
author: "Di Y."
date: "9/2/2020"
output:
  html_document: default
  pdf_document: default
---
# Severe weather events and their conscequences
## Synopsis
In the USA, tornadoes have killed and hurt the largest number of American people.From 1950 to 2011, 5633 people have lost their lives due to tornadoes and more than 91000 people got injured. 
Moreover, tornadoes also caused the largest damage of properties. In the same period, the loss of property was estimated as $90 billion, while hail caused estimatedly total crop damage of more than $4 million. The estimation can be more precise if the type of the weather events was more generalized.

## 1. Data Reading and Processing
```{r, cache=T}
file <- "repdata_data_StormData.csv.bz2"
if (!file.exists(file)) {
  download.file('https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2', file,method="curl")
        unzip("repdata_data_StormData.csv.bz2",exdir="data")
}

raw <- read.csv(file,header=T,sep=",")
dim(raw)
str(raw)
```

```{r,cache=T}
# Let the unit of the damage value in columns "PROPDMGEXP" and "CROPDMGEXP" just be $
summary(raw)
table(raw$PROPDMGEXP)
table(raw$CROPDMGEXP)
unit_fun <- function(unit){
  if (unit=="") {
    return(1)
  } else if(unit %in% c("?","-","+")){
    return(0)
  } else if(as.numeric(unit) %in%c(1:10)){
    return(as.numeric(unit))
  } else if(unit %in% c("k","K")){
    return(10^3)
  } else if(unit %in% c("m","M")){
    return(10^6) 
  } else if(unit %in% c("h","H")){
    return(10^2)
  } else if(unit %in% c("B","b")){
    return(10^9)
  }
    
  }
temp_p <- sapply(raw$PROPDMGEXP,unit_fun)
temp_c <- sapply(raw$CROPDMGEXP,unit_fun)
library(dplyr)
raw <- raw %>% 
  transform(PROPDMG=PROPDMG*unlist(temp_p),CROPDMG=CROPDMG*unlist(temp_c)) # unit is now $
raw <-  raw %>% mutate(PROPDMG_BIO=PROPDMG/10^9,CROPDMG_BIO=CROPDMG/10^9) # unit is now billion $

head(raw)
dim(raw)
```

## Results
### 1) Across the United States, which types of events (as indicated in the EVTYPE\color{red}{\verb|EVTYPE|}EVTYPE variable) are most harmful with respect to population health?
```{r, cache=T}

injury <- raw %>% group_by(EVTYPE) %>%
        summarise(sum_injury=sum(INJURIES,na.rm=T)) %>% 
        arrange(desc(sum_injury))
        

# Top 10 severe weathers, which brought the most injuries
injury_top10 <- injury %>% top_n(10)

fatality <- raw %>% group_by(EVTYPE) %>%
        summarise(sum_fatality=sum(FATALITIES,na.rm=T)) %>% 
        arrange(desc(sum_fatality)) 
        

# Top 10 severe weathers, which led to the most fatalities
fatality_top10 <- fatality %>% top_n(10)
```

```{r, echo=F,cache=T}
library(ggplot2)
install.packages("ggpubr",repos = 'http://cran.us.r-project.org')
library(ggpubr)
```

```{r,echo=F,cache=T}
plot_injury <- injury_top10 %>%
        ggplot(aes(x=reorder(EVTYPE,-desc(sum_injury),sum),y=sum_injury,fill=interaction(EVTYPE,sum_injury)))+
        geom_col(show.legend = F)+
        coord_flip()+
        labs(title="Number of Injuries due to Extreme Weathers",x="Type of Weathers",y="Number of Injuries")+
        geom_text(aes(label=sum_injury),hjust=-0.05,size=2.5)
plot_fatality <- fatality_top10 %>% 
        ggplot(aes(x=reorder(EVTYPE,-desc(sum_fatality),sum),y=sum_fatality,fill=interaction(EVTYPE,sum_fatality)))+
        geom_col(show.legend = F)+
        coord_flip()+
        labs(title="Number of Fatalities due to Extreme Weathers",x="Type of Weathers",y="Number of Fatalities")+
        geom_text(aes(label=sum_fatality),hjust=-0.05,size=2.5)
ggarrange(plot_injury,plot_fatality,labels=c("A","B"),ncol=1,nrow=2)
```

### 2) Across the United States, which types of events have the greatest economic consequences?

```{r,cache=T}
prop_damage <- raw %>% group_by(EVTYPE) %>%
        summarise(damage_p=sum(PROPDMG_BIO,na.rm=T))%>%
        arrange(desc(damage_p)) 
      
prop_damage_top10 <- prop_damage %>% top_n(10)
crop_damage <- raw %>% group_by(EVTYPE) %>%
        summarise(damage_c=sum(CROPDMG_BIO,na.rm=T)) %>%
        arrange(desc(damage_c)) 
        
crop_damage_top10 <- crop_damage %>% top_n(10)
```

```{r, echo=F,cache=T}
prop_dmg <- prop_damage_top10 %>%
        ggplot(aes(x=reorder(EVTYPE,-desc(damage_p)),y=damage_p,fill=interaction(EVTYPE,damage_p)),label=sprintf("%0.2f", round(damage_p, digits = 2)))+
                       geom_col(show.legend = F)+
                       coord_flip()+
                labs(title="Damage of Proporty due to Extreme Weathers",x="Type of Weathers",y="Value of Damaged Proporty (billion $)")
                
crop_dmg <- crop_damage_top10 %>%
        ggplot(aes(x=reorder(EVTYPE,-desc(damage_c)),y=damage_c,fill=interaction(EVTYPE,damage_c)))+
        geom_col(show.legend = F)+
        coord_flip()+
        labs(title="Damage of Crop Damage due to Extreme Weathers",x="Type of Weathers",y="Value of Damanaged Crop  (billion $)")
        
ggarrange(prop_dmg,crop_dmg,labels=c("C","D"),ncol=1,nrow=2)
```
